---
title: "Covid91 vaccine study Final2024"
author: "Hayden Wood!"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

$H_0$ The new preventative vaccine (covid921) does exhibit different effectiveness across the variety of subpopulations tracked

$H_A$ The new preventative vaccine (covid921) does NOT exhibit different effectiveness across the variety of subpopulations tracked

#Methods

Covid921 is a Categorical data set and also Sub-populations is a categorical data set, so we will use the Cat_Cat helper file which will give us codes to help properly decode the dataset we have been provided with. We are going to use a ggplot, tables and fishers tests for each plot of data to help us get a more defined look at the covid921 vaccine's effects on these subcatories. We will also compare the P value for each set of data to see if there are any big changes in these different subcatagories.

## Males

## Descriptive Results

Here we are going to compare the effectiveness in males that have been given the Covid921 using GGplots and filtering bewtween only males


### Graphical Descriptive Results

```{r}
levels(FinalData$sex)
dataG = filter(FinalData,sex=="M")

dd2 <- dataG %>% group_by(treatment,infected) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=treatment,y=count,fill=infected))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=treatment,y=prcnt*100,fill=infected)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This graph suggests that the difference in males that have been given the "placebo" compared to the real "vaccine" has a appeared to be higher when compared to the "vaccine". But not a substantial difference

### Numerical Descriptive Results
```{r}
levels(FinalData$sex)
dataG = filter(FinalData,sex=="M")

table1 <- xtabs(~infected +treatment , data=dataG)
rowPerc(table1)
colPerc(table1)
```
This table suggest that the number of males "M" that recieved the "treatment" has much lower infection rate than those who did not recieve the "treatment". But again not at a substantial rate

## Inferential Results

```{r}
levels(FinalData$sex)
dataG = filter(FinalData,sex=="M")


chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

This ttest statistic shows that the p-value comes out to be  0.00000000000004631 for the effectiveness in males and chi sq shows 0.0000000000001098

## Females

## Descriptive Results

Here we are going to compare the effectiveness in females that have been given the Covid921. With all the codes in this section we make sure to start with a code that filters to make sure we are only using datasets from the gender "f"

### Graphical Descriptive Results
```{r}
levels(dataG$sex)
dataG = filter(FinalData,sex=="F")

dd2 <- dataG %>% group_by(treatment,infected) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=treatment,y=count,fill=infected))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=treatment,y=prcnt*100,fill=infected)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This graph shows the difference between infected patients who received either a placebo vaccine or the actual vaccine between, the gender "f" in females. We make sure to include a code at the begging to make sure we are only using data set "f" in the sex column. This also shows 2 graphs, one in percentages and one in count.

### Numerical Descriptive Results

```{r}
levels(FinalData$sex)
dataG = filter(FinalData,sex=="F")

table1 <- xtabs(~infected +treatment , data=dataG)
rowPerc(table1)
colPerc(table1)
```

This table shows the differences in infected patients that are female that have received that vaccine or placebo vs those who haven't. The percentages suggest that the covid treatment could have a positive affect in showing resistance to the covid disease

## Inferential Results

```{r}
levels(FinalData$sex)
dataG = filter(FinalData,sex=="F")


chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The P value for The gender F for the effectiveness of the vaccine is 0.000000000232 for the T test and 0.000000002842 for the chisq

## LGBTQ

## Descriptive Results

Like we have done previously. We are going to filter the results so it only applies to those appart of the LGBTQ community to see the effectiveness of the covid921 vaccine 

### Graphical Descriptive Results

```{r}
levels(dataG$LGBTQ)
dataG = filter(FinalData,LGBTQ=="gay")

dd2 <- dataG %>% group_by(treatment,infected) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=treatment,y=count,fill=infected))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=treatment,y=prcnt*100,fill=infected)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This graph shows the number of people who revieved the placebo vs the vaccine between people who identify as "gay" that are apart of the LGBTQ. Something interesting that we havent seen in the other graphs is that the graph appears to suggest that that more people that recieved that vaccine in this community has apeared to have accured the virus compared to the placebo. When we conducted the same test of the F for Female and M in male we saw and opposite suggestion. But this difference is still not big enough to make a substantial impact

### Numerical Descriptive Results

```{r}
levels(dataG$LGBTQ)
dataG = filter(FinalData,LGBTQ=="gay")

table1 <- xtabs(~infected +treatment , data=dataG)
rowPerc(table1)
colPerc(table1)
```

this shows percentages of the vaccine and and the placebo dose between people who identify as "gay" in the LBGTQ subcategory. Like shown in the graph,this table suggests we can see that there is a small percentage increase in "gay" and Infected under the vaccine

## Inferential Results

```{r}
levels(dataG$LGBTQ)
dataG = filter(dataG,LGBTQ=="gay")


chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


With the data above and the test we have just conducted. we have shown that the P value is 1.029e-08 for the ttest and 6.668e-08
for the chisq

## Druggies


## Descriptive Results

In this section we are going to take the subcategory of drug users "druggies" and filter their results to see the effectiveness of the Covid vaccine

### Graphical Descriptive Results


```{r}
levels(dataG$DrugUser)
dataG = filter(FinalData,DrugUser=="yes")

dd2 <- dataG %>% group_by(treatment,infected) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=treatment,y=count,fill=infected))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=treatment,y=prcnt*100,fill=infected)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This graph shows the differences between the vaccine and the placebo vaccine between drugusers. We used a code at the begging to filter the category drug users and the "yes" subcategory. This graph suggests that there is a slightly higher infection between drug users that got the placebo to the vaccine.

### Numerical Descriptive Results

```{r}
levels(FinalDat$DrugUser)
FinalDat = filter(FinalDat,DrugUser=="yes")

table1 <- xtabs(~infected +treatment , data=dataG)
rowPerc(table1)
colPerc(table1)
```

This shows differences in drugUsers that use drugs that have been given the placebo and the real vaccine. 

## Inferential Results

```{r}
levels(FinalData$DrugUser)
FinalData = filter(FinalData,DrugUser=="yes")


chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The P value for the people that have been given the to the subcategory "yes" for DrugUsers 2.32e-10 for the ttest and 2.842e-09
for the Chisq

# Overall Results and Conclusions            
Based on the analysis conducted across four different subpopulations (males, females, LGBTQ individuals, and drug users), we have obtained the following p-values, ive simplified it for only the P values from the Ttest because the Chisq supported the ttest in each subcatagory

For males:        
ð‘= 0.00000000000004631
 
For females:         
ð‘= 0.0002732

For LGBTQ individuals:         
ð‘= 0.00000001029
 
For drug users:         
ð‘=0.000000000232

 
Given the significance level typically set at ð›¼= 0.05 we compare these p-values to determine if we should reject the null hypothesis $H_0$ that the new preventative vaccine (covid921) exhibits different effectiveness across the variety of subpopulations tracked.

Considering all four p-values, we observe that each one is substantially smaller than the level of 0.05. In other words, the probability of observing such extreme results under the assumption that the vaccine exhibits different effectiveness across subpopulations is extremely low.

Therefore, based on the statistical analysis conducted, we reject the null hypothesis $H_0$ and conclude that the new preventative vaccine (covid921) does NOT exhibit different effectiveness across the variety of subpopulations tracked. Instead, the evidence suggests that the vaccine's effectiveness is consistent across different demographic subgroups.


